---
title: "mvDFA"
subtitle: "multivariate Detrended Fluctuation Analysis using Package mvDFA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mvDFA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an `R` package to estimate the univariate and a **m**ulti**v**ariate adaption of the Hurst exponent using **D**etrended **F**luctuation **A**nalysis. The package was created as a supplement to the publication **INSERT PUBLICATION HERE: Wallot et al. (2023)**.

## Install the latest working version from Github
This requires the package `devtools`. The following code installs `devtools`, if it is not installed.

```{r, eval=FALSE}
if(!"devtools" %in% installed.packages()) install.packages("devtools")
devtools::install_github("jpirmer/mvDFA", build_vignettes = T)
```

Use `build_vignettes = T` to be able to see the documentation linked in "Getting Started". 

Load the package:

```{r setup}
library(mvDFA)
```


## Test functions for approximated correlated Time Series with different Hurst Exponent

### Simulate correlated time series

Choose a covariance matrix $\Sigma$
```{r}
Sigma <- Sigma <- matrix(.5, 4, 4); diag(Sigma) <- 1
Sigma[3,4] <- Sigma[4,3] <- -.3
Sigma
```
Simulate from the covariance matrix with approximate univariate Hurst-exponents $H:=(.2, .5, .7, .9)'$. 

```{r}
set.seed(2023)
X <- simulate_cMTS(N = 10^3, H = c(.2, .5, .7, .9), Sigma = Sigma, simulation_process = "FGN0", decomposition = "chol")
head(X)
```

`simulation_process = "FGN0"` uses `longmemo::simFGN0` to generate independent (Fractional) Gaussian Processes, which are then mixed using the Cholesky decomposition $D$ of $\Sigma := DD'$. Changing this argument to `simulation_process = "FGN.fft"` uses `longmemo::simFGN.fft` (using FFT) to generate independent (Fractional) Gaussian Processes, which are then mixed using the Cholesky decomposition $D$ of $\Sigma := DD'$. The differences may be inspected here:

```{r, fig.align='center', fig.height=3, fig.width=4}
x1 <- simulate_cMTS(N = 3*10^2, H = c(.5), Sigma = as.matrix(1), simulation_process = "FGN0")
plot(x1$X1, main = "H = 0.5 and FGN0", type = "l")
x2 <- simulate_cMTS(N = 3*10^2, H = c(.5), Sigma = as.matrix(1), simulation_process = "FGN.fft")
plot(x2$X1, main = "H = 0.5 and FGN.fft", type = "l")
```

Note: the true Hurst Exponents might be slightly off, since the mixing (weighted sums) of the time series makes them slightly more normal (due to the central limits theorem). Hence, small Hurst-exponents ($H<0.5$) tend to be upwards biased, while larger Hurst-exponents tend to be downward biased ($H>0.5$). 

Further, the use of Cholesky might influence the generation of data in a different way than another decomposition of $\Sigma$ would. 


### Estimate multivariate extensions of Hurst exponent for correlated time series

The `mvDFA` function needs only the time series in long format as a `matrix` or `data.frame` object. Multiple `cores` can be used (maximum is `parallel::detectCores()`), which reduces computation time drastically for longer timer series. The `steps` argument manipulates the number of window sizes, which are seperated logarithmically. Larger numbers of steps result in more precise estimates of the extended Hurst exponents.

```{r, message=FALSE, warning=FALSE}
mvDFA_result <- mvDFA(X = X, steps = 50, cores = 1)
mvDFA_result
```






```{r,eval=FALSE}

par(mfrow=c(2,3))
plot(out$time, out$X, type = "l", main = "X vs. time")
plot(out$time, out$Y, type = "l", main = "Y vs. time")
plot(out$time, out$Z, type = "l", main = "Y vs. time")
plot(out$X, out$Y, type = "l", main = "Y vs. X")
plot(out$X, out$Z, type = "l", main = "Z vs. X")
plot(out$Y, out$Z, type = "l", main = "Z vs. Y")
```
