---
title: "multivariate Detrended Fluctuation Analysis using Package mvDFA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an `R` package to estimate the univariate and a **m**ulti**v**ariate adaption of the Hurst exponent using **D**etrended **F**luctuation **A**nalysis. The package was created as a supplement to the publication **INSERT PUBLICATION HERE: Wallot et al. (2023)**.

## Install the latest working version from Github
This requires the package `devtools`.

```{r, eval=FALSE}
install.packages("devtools")
devtools::install_github("jpirmer/mvDFA")
```

Load the package:

```{r setup}
library(mvDFA)
```


## Test functions for approximated correlated Time Series with different Hurst Exponent

### Simulate correlated time series
Choose a covariance matrix $\Sigma$
```{r}
Sigma <- Sigma <- matrix(.5, 4, 4); diag(Sigma) <- 1
Sigma
```
Simulate from the covariance matrix with approximate univariate Hurst-exponents $H:=(.2, .5, .7, .9)'$. 

```{r}
set.seed(2023)
X <- simulate_cMTS(N = 10^3, H = c(.2, .5, .7, .9), Sigma = Sigma)
head(X)
```

Note: the true Hurst Exponents might be slightly off, since the mixing (weighted sums) of the time series makes them slightly more normal. Hence, small Hurst-exponents ($H<0.5$) tend to be upwards biased, while larger Hurst-exponents tend to be downward biased ($H>0.5$). 


### Estimate multivariate extensions of Hurst exponent for correlated time series

The `mvDFA` function needs only the time series in long format as a `matrix` or `data.frame` object. Multiple `cores` can be used (maximum is `parallel::detectCores()`), which reduces computation time drastically for longer timer series. The `steps` argument manipulates the number of window sizes, which are seperated logarithmically. Larger numbers of steps result in more precise estimates of the extended Hurst exponents.

```{r, message=FALSE, warning=FALSE}
mvDFA_result <- mvDFA(X = X, steps = 50, cores = 1)
mvDFA_result
```
